Рекурсивные свойства ветвления и БГГ резольвента

### 2011-02-12 AffineLieAlgebras/MQFT2010/article-ru.tex
More translation

\begin{statement}
\bigskip For an orthogonal subalgebra $\frak{a}_{\perp }$ in $\frak{g}$ (an
orthogonal partner of a reductive $\frak{a}\hookrightarrow \frak{g}$) the
character of an integrable highest weight module $L^{\mu }$ can be presented
as a combination (with integral coefficients) of parabolic Verma modules
distributed by the set of weights $e^{\mu _{\widetilde{\mathfrak{a}}}\left(
u\right) }$:
\begin{equation}
\mathrm{ch}\left( L^{\mu }\right) =\sum_{u\in U}\;\epsilon (u)e^{\mu _{%
\widetilde{\frak{a}}}\left( u\right) }\mathrm{ch}M_{I}^{\mu _{\frak{a}%
_{\perp }}\left( u\right) },  \label{gen Weyl-Verma}
\end{equation}
where $U:=\left\{ u\in W|\quad \mu _{\frak{a}_{\perp }}\left( u\right) \in
\overline{C_{\frak{a}_{\perp }}}\right\} $ and $I$ is such a subset of $S$ that
$\Delta _{I}^{+}$ is
equivalent to $\Delta _{\frak{a}_{\perp }}^{+}$.
\end{statement}

\bigskip
\begin{proof}
By the definition (\ref{delta a ort}) the subalgebra $%
\mathfrak{a}_{\bot }$ is regular and reductive. Consider its Weyl
denominator $R_{\frak{a}_{\perp }}:=\prod_{\alpha \in \Delta _{\frak{a}%
_{\perp }}^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{mult}_{\frak{a}}%
\mathrm{\left( \alpha \right) }}$ and the element $R_{J}:=\prod_{\alpha \in
\Delta ^{+}\setminus \Delta _{\frak{a}_{\perp }}^{+}}\left( 1-e^{-\alpha
}\right) ^{\mathrm{mult}(\alpha )}$ as the factors in $R$: $\quad $%
\begin{equation*}
R=R_{J}R_{\frak{a}_{\perp }}.
\end{equation*}
According to this factorization and the decomposition (\ref{sing decomp main}%
) the character $\mathrm{ch}\left( L^{\mu }\right) $ can be written as
\begin{eqnarray*}
\mathrm{ch}\left( L^{\mu }\right) &=&\left( R_{J}\right) ^{-1}\left( R_{%
\frak{a}_{\perp }}\right) ^{-1}\Psi ^{\mu }=\left( R_{J}\right)
^{-1}\sum_{u\in U}\;e^{\mu _{\widetilde{\frak{a}}}\left( u\right) }\epsilon
(u)\left( R_{\frak{a}_{\perp }}\right) ^{-1}\Psi _{\frak{a}_{\perp }}^{\mu _{%
\frak{a}_{\perp }}\left( u\right) } \\
&=&\left( R_{J}\right) ^{-1}\sum_{u\in U}\;e^{\mu _{\widetilde{\frak{a}}%
}\left( u\right) }\epsilon (u)\mathrm{ch}\left( L_{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp
}}\left( u\right) }\right),
\end{eqnarray*}
where $\left\{ L_{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp }}\left(
u\right) }|u\in U\right\} $ is the set of finite-dimensional $\frak{a}%
_{\perp }$-modules with the highest weights $\mu _{\frak{a}_{\perp }}\left(
u\right) $. We are interested in nontrivial subalgebras $\frak{a}$ and
correspondingly in nontrivial $\frak{a}_{\perp }$ (the case of a trivial
orthogonal subalgebra was considered above (see Remark 1)). This means that $r_{\frak{a}}\geq 1$ and $r_{\frak{a}%
_{\perp }}<r$. Due to the fact that any maximal regular subalgebra has the
Dynkin scheme obtained by one or two node subtractions from the extended
Dynkin scheme and the extended scheme has at most one dependent root (the
highest root) the set of roots $\Delta _{\frak{a}_{\perp }}^{+}$ is always
equivalent to the one $\Delta _{I}^{+}$ generated\ by some subset $I\subset
S $ of simple roots.

It follows that we can (by redefining the set $\Delta ^{+}$) identify $%
\Delta _{\frak{a}_{\perp }}^{+}$ with the subset $\Delta _{I}^{+}$ \ where $%
I\subset S$ . This allows us to introduce the elements necessary to compose
the generalized Verma modules
\cite{lepowsky1977generalization,humphreys2008representations}.
We have two sets of root
vectors $\left\{ x_{\xi }\in \frak{g}_{\xi }|\xi \in \Delta _{I}^{+}\right\}
$ and $\left\{ x_{\eta }\in \frak{g}_{\eta }|\eta \in \Delta ^{+}\setminus
\Delta _{I}^{+}\right\} $ and the corresponding nilpotent subalgebras in $\frak{n}%
^{+}$:
\begin{equation*}
\frak{n}_{I}^{+}:=\sum_{\xi \in \Delta _{I}^{+}}\frak{g}_{\xi },\quad
\frak{u}_{I}^{+}:=\sum_{\eta \in \Delta ^{+}\setminus \Delta _{I}^{+}}\frak{g}%
_{\eta }.
\end{equation*}
The first subalgebra together with its negative counterpart $\frak{n}%
_{I}^{-} $ generates a simple subalgebra
\begin{equation*}
\frak{s}_{I}=\frak{n}_{I}^{-}+\frak{h}_{I}+\frak{n}_{I}^{+}.
\end{equation*}
We enlarge it with the remaining Cartan generators:
\begin{equation*}
\frak{l}_{I}=\frak{n}_{I}^{-}+\frak{h}+\frak{n}_{I}^{+}.
\end{equation*}
The semidirect product of $\frak{l}_{I}$ and $\frak{u}_{I}^{+}$
gives a parabolic subalgebra $\frak{p}_{I}\hookrightarrow \frak{g}$ :
\begin{equation}
\frak{p}_{I}=\frak{l}_{I}\vartriangleright \frak{u}_{I}^{+}.
\label{paralolic subalg}
\end{equation}
Its universal enveloping $U\left( \frak{p}_{I}\right) $ is a subalgebra in $%
U\left( \frak{g}\right) $. The $\frak{l}_{I}$-modules $L_{\frak{a}_{\perp }}^{\mu _{%
\frak{a}_{\perp }}\left( u\right) }$ can be easily lifted to $\frak{p}_{I}$%
-modules using the trivial action of the nilradical $\frak{u}_{I}^{+}$. The
latter induce $U\left( \frak{g}\right) $-modules in a standard way:
\begin{equation*}
M_{I}^{\mu _{\frak{a}_{\perp }}\left( u\right) }=U\left( \frak{g}\right)
\otimes _{U\left( \frak{p}_{I}\right) }L_{\frak{a}_{\perp }}^{\mu _{\frak{a}%
_{\perp }}\left( u\right) }.
\end{equation*}

These are the \textit{generalized Verma modules}  \cite{lepowsky1977generalization}
generated by the highest weights $\mu _{\frak{a}%
_{\perp }}\left( u\right) $. As a $U\left( \frak{u}_{I}^{-}\right) $-module
each $M_{I}^{\mu _{\frak{a}_{\perp }}\left( u\right) }$ is isomorphic to $%
U\left( \frak{u}_{I}^{-}\right) \otimes $ $L_{\frak{a}_{\perp }}^{\mu _{%
\frak{a}_{\perp }}\left( u\right) }$ and thus its character can be written
in terms of Kostant-Heckman function \cite{KostantHeckman1982} corresponding
to the injection of the orthogonal partner $\frak{a}_{\perp
}\hookrightarrow \frak{g}$:
\begin{equation*}
\mathrm{ch}M_{I}^{\mu _{\frak{a}_{\perp }}\left( u\right) }=\mathcal{KH}_{%
\frak{a}_{\perp }\hookrightarrow \frak{g}}\mathrm{ch}L_{\frak{a}_{\perp
}}^{\mu _{\frak{a}_{\perp }}\left( u\right) }.
\end{equation*}
The function $\mathcal{KH}_{\frak{a}_{\perp }\hookrightarrow \frak{%
g}}$ is generated by the denominator $R_{I}$ thus the last expression can be
written in the form
\begin{equation*}
\mathrm{ch}M_{I}^{\mu _{\frak{a}_{\perp }}\left( u\right) }=\frac{1}{R_{I}}%
\mathrm{ch}L_{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp }}\left( u\right) }.
\end{equation*}
This means that we have obtained the generalized Weyl-Verma character formula
 -- the decomposition of $\mathrm{ch}\left(
L^{\mu }\right) $ in terms of generalized Verma module characters:
\begin{equation}
\mathrm{ch}\left( L^{\mu }\right) =\sum_{u\in U}\;e^{\mu _{\aft}\left(
u\right) }\epsilon (u)\mathrm{ch}M_{I}^{\mu _{\frak{a}_{\perp }}\left(
u\right) }.  \label{char in gen verma mod}
\end{equation}
\end{proof}

>>>>

\begin{statement}
\bigskip
Для ортогональной подалгебры  $\frak{a}_{\perp }$ в $\frak{g}$ (являющейся ортогональным партнером редуктивной подалгебры $\frak{a}\hookrightarrow \frak{g}$) характера интегрируемого модуля старшего веса  $L^{\mu }$ может быть представлен в виде комбинации (с целочисленными коэффициентами) параболических модулей Верма, распределенных по множеству весов $e^{\mu _{\widetilde{\mathfrak{a}}}\left(
u\right) }$:
\begin{equation}
\mathrm{ch}\left( L^{\mu }\right) =\sum_{u\in U}\;\epsilon (u)e^{\mu _{%
\widetilde{\frak{a}}}\left( u\right) }\mathrm{ch}M_{I}^{\mu _{\frak{a}%
_{\perp }}\left( u\right) },  \label{gen Weyl-Verma}
\end{equation}
где  $U:=\left\{ u\in W|\quad \mu _{\frak{a}_{\perp }}\left( u\right) \in
\overline{C_{\frak{a}_{\perp }}}\right\} $ и $I$ -- такое подмножество в  $S$, что $\Delta _{I}^{+}$ эквивалентно $\Delta _{\frak{a}_{\perp }}^{+}$.
\end{statement}

\bigskip
\begin{proof}
Подалгебра  $\mathfrak{a}_{\bot }$ регулярна и редуктивна по определению  (\ref{delta a ort}). Рассмотрим её знаменатель Вейля   $R_{\frak{a}_{\perp }}:=\prod_{\alpha \in \Delta _{\frak{a}%
_{\perp }}^{+}}\left( 1-e^{-\alpha }\right) ^{\mathrm{mult}_{\frak{a}}%
\mathrm{\left( \alpha \right) }}$ и элемент  $R_{J}:=\prod_{\alpha \in
\Delta ^{+}\setminus \Delta _{\frak{a}_{\perp }}^{+}}\left( 1-e^{-\alpha
}\right) ^{\mathrm{mult}(\alpha )}$ как сомножители в $R$:
\begin{equation*}
R=R_{J}R_{\frak{a}_{\perp }}.
\end{equation*}
Согласно этой факторизации и разложению  (\ref{sing decomp main}) характер $\mathrm{ch}\left( L^{\mu }\right) $ можно переписать в виде
\begin{eqnarray*}
\mathrm{ch}\left( L^{\mu }\right) &=&\left( R_{J}\right) ^{-1}\left( R_{%
\frak{a}_{\perp }}\right) ^{-1}\Psi ^{\mu }=\left( R_{J}\right)
^{-1}\sum_{u\in U}\;e^{\mu _{\widetilde{\frak{a}}}\left( u\right) }\epsilon
(u)\left( R_{\frak{a}_{\perp }}\right) ^{-1}\Psi _{\frak{a}_{\perp }}^{\mu _{%
\frak{a}_{\perp }}\left( u\right) } \\
&=&\left( R_{J}\right) ^{-1}\sum_{u\in U}\;e^{\mu _{\widetilde{\frak{a}}%
}\left( u\right) }\epsilon (u)\mathrm{ch}\left( L_{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp
}}\left( u\right) }\right),
\end{eqnarray*}
где  $\left\{ L_{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp }}\left(
u\right) }|u\in U\right\} $ -- множество конечномерных $\frak{a}%
_{\perp }$-модулей со старшими весами   $\mu _{\frak{a}_{\perp }}\left(
u\right) $. Нас интересуют нетривиальные подалгебры $\frak{a}$ и, соответственно, нетривиальные  $\frak{a}_{\perp }$ (случай тривиальной ортогональной подалгебры рассматривался выше в (см. Замечание 1)). Значит $r_{\frak{a}}\geq 1$ и $r_{\frak{a}%
_{\perp }}<r$. Так как диаграмма Дынкина для любой регулярной подалгебры получается в результате исключения одной, двух или более вершин из расширенной диаграммы Дынкина алгебры, а расширенная диаграмма содержит не более одного зависимого корня (старший корень), множество корней $\Delta _{\frak{a}_{\perp }}^{+}$ всегда эквивалентно некоторому множеству $\Delta _{I}^{+}$, порожденному набором  простых корней $I\subset S $.

Следовательно мы можем (переопределяя множество  $\Delta ^{+}$) отождествить $\Delta _{\frak{a}_{\perp }}^{+}$ с подмножеством  $\Delta _{I}^{+}$, где $I\subset S$. В результате мы можем определить объекты, необходимые для построения обобщенных модулей Верма \cite{lepowsky1977generalization,humphreys2008representations}. У нас есть два множества корневых векторов $\left\{ x_{\xi }\in \frak{g}_{\xi }|\xi \in \Delta _{I}^{+}\right\}
$ и $\left\{ x_{\eta }\in \frak{g}_{\eta }|\eta \in \Delta ^{+}\setminus
\Delta _{I}^{+}\right\} $, и соответствующие нильпотентные подалгбры в  $\frak{n}%
^{+}$:
\begin{equation*}
\frak{n}_{I}^{+}:=\sum_{\xi \in \Delta _{I}^{+}}\frak{g}_{\xi },\quad
\frak{u}_{I}^{+}:=\sum_{\eta \in \Delta ^{+}\setminus \Delta _{I}^{+}}\frak{g}%
_{\eta }.
\end{equation*}
Первая подалгебра, вместе со своей отрицательной копией  $\frak{n}%
_{I}^{-} $ порождает простую подалгебру 
\begin{equation*}
\frak{s}_{I}=\frak{n}_{I}^{-}+\frak{h}_{I}+\frak{n}_{I}^{+}.
\end{equation*}
Мы расширяем ее оставшимися картановскими генераторами:
\begin{equation*}
\frak{l}_{I}=\frak{n}_{I}^{-}+\frak{h}+\frak{n}_{I}^{+}.
\end{equation*}
Полупрямое произведение  $\frak{l}_{I}$ и $\frak{u}_{I}^{+}$ дает параболическую подалгебры $\frak{p}_{I}\hookrightarrow \frak{g}$ :
\begin{equation}
\frak{p}_{I}=\frak{l}_{I}\vartriangleright \frak{u}_{I}^{+}.
\label{paralolic subalg}
\end{equation}
Ее универсальная обертывающая $U\left( \frak{p}_{I}\right) $ является подалгеброй в  $%
U\left( \frak{g}\right) $. Модули  $L_{\frak{a}_{\perp }}^{\mu _{%
\frak{a}_{\perp }}\left( u\right) }$ алгебры  $\frak{l}_{I}$ легко поднимаются до  $\frak{p}_{I}$-модулей при помощи тривиального действия нильрадикала $\frak{u}_{I}^{+}$. Последний стандартным образом индуцирует $U\left( \frak{g}\right) $-модули:
\begin{equation*}
M_{I}^{\mu _{\frak{a}_{\perp }}\left( u\right) }=U\left( \frak{g}\right)
\otimes _{U\left( \frak{p}_{I}\right) }L_{\frak{a}_{\perp }}^{\mu _{\frak{a}%
_{\perp }}\left( u\right) }.
\end{equation*}

Это  \textit{обобщенные модули Верма}  \cite{lepowsky1977generalization}
порожденные старшими весами  $\mu _{\frak{a}%
_{\perp }}\left( u\right) $. Как  $U\left( \frak{u}_{I}^{-}\right) $-модуль каждый  $M_{I}^{\mu _{\frak{a}_{\perp }}\left( u\right) }$ изоморфен  $%
U\left( \frak{u}_{I}^{-}\right) \otimes L_{\frak{a}_{\perp }}^{\mu _{%
\frak{a}_{\perp }}\left( u\right) }$ и его характер можно записать при помощи функции Костанта-Хекмана \cite{KostantHeckman1982}, соответствующей вложению ортогонального партнера $\frak{a}_{\perp
}\hookrightarrow \frak{g}$:
\begin{equation*}
\mathrm{ch}M_{I}^{\mu _{\frak{a}_{\perp }}\left( u\right) }=\mathcal{KH}_{%
\frak{a}_{\perp }\hookrightarrow \frak{g}}\mathrm{ch}L_{\frak{a}_{\perp
}}^{\mu _{\frak{a}_{\perp }}\left( u\right) }.
\end{equation*}
Функция  $\mathcal{KH}_{\frak{a}_{\perp }\hookrightarrow \frak{%
g}}$ генерируется знаменателем  $R_{I}$, так что последнее выражение можно переписать следующим образом
\begin{equation*}
\mathrm{ch}M_{I}^{\mu _{\frak{a}_{\perp }}\left( u\right) }=\frac{1}{R_{I}}%
\mathrm{ch}L_{\frak{a}_{\perp }}^{\mu _{\frak{a}_{\perp }}\left( u\right) }.
\end{equation*}
Значит мы получили обобщенную формулу Вейля-Верма для характеров -- разложение  $\mathrm{ch}\left(
L^{\mu }\right) $ на характеры обобщенных модулей Верма:
\begin{equation}
\mathrm{ch}\left( L^{\mu }\right) =\sum_{u\in U}\;e^{\mu _{\aft}\left(
u\right) }\epsilon (u)\mathrm{ch}M_{I}^{\mu _{\frak{a}_{\perp }}\left(
u\right) }.  \label{char in gen verma mod}
\end{equation}
\end{proof}
